\chapter{Background}
\label{chapter:background}

This chapter covers the concepts the rest of the thesis builds on: what QUIC
is and why it exists, the five implementations under test, the network
parameters used to describe a connection, how those conditions are emulated
with Linux traffic control, and the machine learning method behind the
recommender.

% =============================================================================
\section{The QUIC Protocol}
\label{sec:bg-quic}

QUIC is a transport protocol that runs on top of UDP. It started at Google
and was later standardised by the IETF in RFC~9000~\cite{rfc9000}. The reason
it exists is that TCP combined with TLS has a few problems that are hard to
fix without breaking compatibility with the existing internet.

The first is \textbf{head-of-line blocking}. TCP delivers a single ordered byte
stream, so if one packet is lost, everything behind it waits until that packet
is retransmitted, even when those later bytes belong to a completely unrelated
request. QUIC avoids this by carrying multiple independent streams in one
connection: a loss on one stream does not stall the others.

The second is \textbf{handshake latency}. A TCP connection needs its own handshake, and
then TLS needs another one on top. QUIC folds the transport and cryptographic
handshakes together~\cite{rfc9001}, so a new connection is ready after one
round trip, or zero round trips if the client has talked to the server
before.

The third is \textbf{ossification}. Because TCP lives in the operating system kernel
and its headers are visible on the wire, middleboxes along the path came to
depend on its exact behaviour, which makes the protocol very hard to change.
QUIC runs in user space and encrypts almost everything, including most of its
own control information, so it can keep evolving without middleboxes getting
in the way. Running in user space has a second benefit for mobile clients:
QUIC connections are identified by a \emph{connection ID} rather than the usual
four-tuple (source IP, source port, destination IP, destination port), so a connection
can survive the client changing its IP address, for example when a phone moves from
WiFi to cellular.

\textbf{Loss detection and congestion control} are specified separately in
RFC~9002~\cite{rfc9002}, and the version-independent properties of the wire
format are described in RFC~8999~\cite{rfc8999}. The original design and
large-scale deployment experience are documented by Langley et
al.~\cite{langley2017quic}.

QUIC is not a free win, though. Running the transport in user space means more
CPU work per byte than the kernel TCP stack, and on very fast links this
overhead can actually make QUIC slower than TCP. Recent work by Zhang et
al.~\cite{zhang2024quic} measured the UDP+QUIC+HTTP/3 stack losing up to
45\% of throughput compared to TCP+TLS+HTTP/2 over fast connections. This
trade-off, where QUIC helps in some conditions and hurts in others, is exactly
what motivates this thesis.

Part of that per-byte cost comes from sending and receiving one UDP datagram at
a time, since each datagram crosses the user/kernel boundary on its own. Linux
offers two offload mechanisms to amortise this. \textbf{Generic Segmentation
Offload (GSO)} lets the application hand the kernel one large buffer plus a
segment size, while the kernel (or the NIC) splits it into MTU-sized datagrams late
in the send path, so a single system call emits many packets. \textbf{Generic
Receive Offload (GRO)} is the receive-side counterpart: it coalesces several
incoming datagrams into one buffer before passing them up. Both cut the number
of system calls per byte, which is exactly the cost that hurts QUIC on fast
links, so they matter for any high-throughput UDP transport. These
abbreviations are used in the rest of the thesis.

% =============================================================================
\section{QUIC Implementations}
\label{sec:bg-implementations}

There is one QUIC specification but many independent implementations of it,
and they do not all behave the same way. Marx et al.~\cite{marx2020same}
studied this directly and found large differences between stacks in how they
handle flow control, congestion control, multiplexing, and other features,
along with bugs and missing optimisations. So "which QUIC" is a real question,
not a minor detail, and it is the question this thesis tries to answer.

This work tests five implementations. For each of the four QUIC stacks we wrote
our own continuous-streaming server and client against the library's API rather
than using any bundled tool, so that all of them share the same workload and
measurement logic (Section~\ref{sec:impl-interop-runner}).

\begin{itemize}
  \item \textbf{quiche} (Cloudflare, written in Rust). Its defining trait is a
        \emph{sans-I/O} design: the library handles QUIC connection state but
        leaves the UDP socket and event loop to the application, so the caller
        decides how packets are sent and received. Tunable parameters such as
        the flow-control windows, idle timeout, and congestion-control
        algorithm are set at runtime through a \texttt{Config} object rather
        than chosen at compile time. Because the socket is the caller's
        responsibility, the high-throughput send path (GSO and kernel pacing)
        has to be wired up in application code. Our server does this following
        Cloudflare's reference examples~\cite{cloudflare-quiche-gso}.
  \item \textbf{msquic} (Microsoft, written in C). A cross-platform core that
        backs HTTP/3 on Windows and in .NET, built around its own worker-thread
        datapath. It becomes flow-control limited at high bandwidth-delay
        products, a detail that matters in the results
        (Section~\ref{sec:bg-network-params}).
  \item \textbf{mvfst} (Meta, written in C++). Built on Meta's Folly and Fizz
        libraries and run in production for Facebook and Instagram traffic.
        Unlike the others it does not expose the socket to the caller. Our
        benchmark server and client are written directly against its
        \texttt{QuicServer}/\texttt{QuicClient} API on top of a Folly event loop.
  \item \textbf{quic-go} (written in Go). A pure-Go implementation with no C
        dependencies, which makes it easy to embed (it backs the Caddy web
        server, among others). It runs entirely in user space on top of Go's
        runtime and garbage collector.
  \item \textbf{TCP with TLS~1.3}, used as the baseline the QUIC
        implementations are compared against. For brevity, this stack is
        referred to as simply \emph{TCP} throughout the thesis.
\end{itemize}

In the configurations measured here, none of these stacks use newer kernel
interfaces such as \texttt{io\_uring}, they all run over the conventional
epoll-based UDP path, batching packets with GSO on the send side. The differences between these stacks, in throughput, CPU usage, and
behaviour under loss, are what the dataset captures and what the recommender
learns from.

% =============================================================================
\section{Network Condition Parameters}
\label{sec:bg-network-params}

A connection is described here by a small set of parameters, each of which is
varied during data collection:

\begin{itemize}
  \item \textbf{Bandwidth} -- the capacity of the link, in Mbps.
  \item \textbf{RTT} -- the round-trip time, how long a packet takes to reach
        the other end and come back.
  \item \textbf{Loss} -- the fraction of packets dropped on the way.
  \item \textbf{Jitter} -- variation in delay from one packet to the next.
  \item \textbf{Corruption} -- the fraction of packets that arrive with
        flipped bits.
  \item \textbf{Competing traffic} -- background TCP or UDP load sharing the
        link.
\end{itemize}

The one concept that ties bandwidth and RTT together is the
\textbf{bandwidth-delay product} (BDP), which is bandwidth multiplied by RTT. It is the amount of data
that has to be in flight at once to keep a link fully busy. The BDP matters a
lot in practice: an implementation whose flow-control window is smaller than
the BDP cannot fill the pipe, no matter how fast the link is. This is the main
reason msquic struggles at high bandwidth and high RTT. In this work msquic
starts each stream with a 32~MB receive window (it can auto-tune higher, but
does so slowly), and once the BDP grows past that window its reliability drops
sharply. The \texttt{fc\_limited} feature used by the model marks conditions
where the BDP exceeds 32~MB, which is where the problem empirically sets in.

% =============================================================================
\section{Network Emulation with tc/netem}
\label{sec:bg-netem}

The network conditions above are not collected from a real network but
emulated on a Linux host, using the kernel's traffic control subsystem
(\texttt{tc}). The central concept in \texttt{tc} is the \textbf{queueing
discipline (qdisc)}: the algorithm attached to a network interface that decides
how outgoing packets are queued, delayed, reordered, or dropped before they
leave. Every interface has a qdisc, which by default is a simple first-in
first-out queue, but it can be replaced with one that actively shapes or impairs
traffic.

This work uses a single \texttt{netem} (network emulation)
qdisc~\cite{hemminger2005netem} per interface. netem injects delay, jitter,
loss, and corruption, and its built-in \texttt{rate} parameter also caps the
bandwidth, so one qdisc both shapes and impairs the link. Emulation is used
instead of a real network because it is repeatable: the exact same conditions
can be applied again and again, which is what makes a clean dataset possible.

With this in mind, and also because it shows up in the results, it is worth noting that
netem rules are applied per interface, on the way out. The simulator in this
setup sits between the client and the server with rules on both of its
interfaces, so a packet making a full round trip crosses the impairment twice.
That is why, in the evaluation, the loss rate measured by the prober comes out
higher than the value that was configured.

% =============================================================================
\section{Random Forest Classifiers}
\label{sec:bg-random-forest}

The recommender is built on random forests. A decision tree splits the data on
feature thresholds, one question at a time, until it reaches a prediction. A
single tree is easy to read but tends to overfit, learning the training data
too closely and generalising poorly. A random forest fixes this by training
many trees, each on a bootstrapped sample of the data and a random subset of
the features, and averaging their predictions~\cite{breiman2001random}. The
averaging cancels out the noise of individual trees and gives a model that
generalises well.

It is worth saying briefly why a random forest, and not something else. A
\emph{neural network} would be overkill: the data here is tabular and fairly small,
on the order of tens of thousands of conditions with eight features, while
deep models want far more data and a lot of tuning, train slowly, and do not
explain themselves. \emph{Linear models} such as logistic regression are the
opposite problem, too weak, because the important effects are non-linear, like
the BDP threshold that decides whether msquic times out or the way loss and
RTT interact. \emph{Gradient boosting} (XGBoost~\cite{chen2016xgboost},
LightGBM~\cite{ke2017lightgbm}) is the honest closest competitor and is often
just as good or better, but it is more sensitive to its hyperparameters,
whereas a random forest gives robust results with very little
tuning~\cite{probst2019hyperparameters}.

A different framing worth mentioning is \emph{online learning}, such as multi-armed
bandits~\cite{vermorel2005multiarmed} or contextual
bandits~\cite{lu2010contextual}, which treat selection as an
explore-and-exploit problem: keep trying options in production and learn from
the outcomes. The exploration part is what does not fit here. The dataset is
collected offline and labelled, so the model can be trained once and then make
a recommendation straight from measured conditions, without ever having to
deploy a deliberately suboptimal implementation just to see how it does. That
matters because a bad choice on a live connection has a real cost. This does
not rule out running the recommender continuously: the prober can re-measure
and re-recommend on a schedule, and a small hysteresis gate keeps it from
flipping between two nearly equal implementations on a noisy link. But every
one of those decisions is still a one-shot supervised prediction, not an
experiment. Supervised classification is the right fit, and random forests
have also been used successfully on QUIC traffic in related
work~\cite{app2023quicensemble}.

Two more features of random forests matter for this thesis. The first is
feature importance: the model can report how much each input contributes to
its decisions, which keeps the recommendations explainable rather than a black
box. This work uses permutation importance, measured on held-out data, rather
than the default impurity-based importance, because the impurity measure is
biased toward high-cardinality continuous features~\cite{strobl2007bias}.

The second is cross-validation. Each network condition in the dataset is
measured ten times to average out noise, which means a plain random
train/test split would put near-identical rows on both sides and make the
model look more accurate than it really is. To avoid this leakage, all
accuracy numbers use GroupKFold cross-validation, which keeps every repetition
of a given condition together in the same fold. This is what makes the
accuracy figures in the evaluation trustworthy.
