\chapter{Implementation}
\label{chapter:implementation}

\section{Initial Approach: quic-interop-runner}
\label{sec:impl-interop-runner}

Initially, we opted to use the QUIC Interop Runner~\cite{quic-interop-runner},
an open-source framework maintained by the QUIC working group for testing
interoperability between QUIC implementations. Each implementation provides a
standardised Docker image exposing a common interface. Meanwhile, the runner orchestrates
client-server pairs through a set of predefined test scenarios executed inside a
network simulator container.

This framework was chosen because it already included Docker images for
several relevant implementations and an integrated network simulation layer,
which appeared to reduce the setup effort for a benchmarking study. However,
two fundamental problems emerged during adaptation.

\subsection{Fixed file transfer model}
\label{subsec:file-transfer-model}

The official runner images are built around a \textit{file transfer} paradigm:
the client requests a file of a fixed size (e.g.\ 10\,MB) and the test ends
when the transfer completes. For interoperability checks, this approach works fine,
but it breaks down for performance benchmarking on high-bandwidth links. At
10\,Gbps, saturating the link for even five seconds requires transferring
approximately 6.25\,GB of data. Extending this to the range of bandwidths
tested in this paper would have required generating and transferring tens to
hundreds of gigabytes per test run, making the evaluation of a big number of
parameter combinations completely impractical.

To address this, we built a set of custom \textit{continuous-streaming}
implementations under \texttt{quic-continuous-implementations/}.
These servers and clients stream data indefinitely for a fixed wall-clock duration,
operating the same way as \texttt{iperf3} does for TCP. A TCP baseline was added alongside
them. We also extended the runner with batch testing infrastructure\
(\texttt{run\_measurements.py}) to go through network condition combinations
automatically.

\subsection{Prohibitive run times and suspicious results}
\label{subsec:interop-problems}

After integrating the continuous implementations and enabling network condition
injection through the runner's ns3-based simulator, two critical problems made
the approach unviable.

\textbf{Run time.} A single scenario (one implementation pair under one network
condition) took between 31 and 62 minutes to complete (measured: TCP 59\,min,
quic-go 50\,min, quiche 46\,min, msquic 31\,min, ngtcp2/nginx 62\,min).
The overhead originated from the ns3 simulation layer, which processes every
packet through a software simulation stack rather than passing it directly to
the OS network stack. In other words, going through a modest condition grid of a
few hundred combinations would have taken weeks of continuous execution.

\textbf{Suspicious results.} Beyond the run time, the measured performance
figures were clearly artifacts of the simulation layer rather than
representative of real network and protocol behaviour. As a concrete example, quic-go
transferred a 10\,MB file over a nominal 10\,Mbps link in 57,962\,ms, nearly
eight times longer than the theoretical minimum of approximately 8\,seconds.
ngtcp2 paired with nginx took 76,688\,ms for the same transfer, an order of
magnitude above the expected duration. These discrepancies were consistent
across runs and across implementations, showing ns3 overhead rather than real
implementation behaviour.

Given these two constraints, the quic-interop-runner approach was abandoned in
favour of a custom testing framework that applies network conditions directly
at the OS level, described in the following section.

% =============================================================================
\section{Custom Testing Framework}
\label{sec:custom-runner}

The custom runner replaces the ns3-based simulation layer with a lightweight
three-container Docker topology. Network impairments are applied
directly by the Linux kernel via \texttt{tc/netem}, which eliminates the
per-packet simulation overhead and produces measurements that reflect
QUIC stack behaviour on Linux.

\subsection{Container topology}
\label{subsec:topology}

Each test run instantiates three containers from a single
\texttt{docker-compose.yml}: \texttt{server}, \texttt{sim}, and
\texttt{client}. The server and client containers each belong to a dedicated
Docker bridge network, \texttt{server\_net} (172.28.100.0/24) and
\texttt{client\_net} (172.28.0.0/24) respectively. The simulator container is
connected to both networks simultaneously and acts as the only forwarding path
between the two endpoints.

Static IP addresses are assigned at container startup: the server receives
172.28.100.100, the client receives 172.28.0.100, and the simulator occupies
172.28.100.2 on the server side and 172.28.0.2 on the client side. Each
endpoint container adds a static route at startup so that traffic destined for
the remote subnet is directed through the simulator.

The implementation that needs to be tested is selected at compose time via
the \texttt{IMPL} environment variable (\texttt{quiche}, \texttt{quic-go},
\texttt{msquic}, \texttt{mvfst}, or \texttt{tcp}), which causes the appropriate
server and client Dockerfiles to be used while the simulator image remains fixed.

\subsection{Network condition injection}
\label{subsec:netem}

All network impairments are applied inside the simulator container using a
two-layer \texttt{tc} configuration. An HTB (Hierarchical Token Bucket) qdisc
enforces the configured bandwidth ceiling. Meanwhile, a \texttt{netem} qdisc attached as
a child leaf introduces delay, jitter, packet loss, and bit corruption.
This configuration is applied independently to both egress interfaces of the
simulator (\texttt{eth0} facing the client network and \texttt{eth1} facing the
server network), so that impairments affect traffic in both directions.

Because of this, a one-way packet traverses exactly one netem stage, giving
a one-way delay equal to \texttt{DELAY}, making the round-trip time equal to
\texttt{2 x DELAY}. Thus, all condition parameters in the dataset are expressed in terms
of RTT, with the netem \texttt{DELAY} value set to \texttt{RTT / 2} accordingly.

Competing background traffic is optionally generated by an \texttt{iperf3}
process inside the simulator container, directed at a dedicated
\texttt{iperf3-target} container on the client network. This traffic, in turn, traverses
the same netem qdisc as the application under test, approximating cross-traffic
contention on a shared bottleneck link.

\subsection{Condition grid}
\label{subsec:condition-grid}

The parameters are defined in \texttt{matrix\_config.sh} and vary across
five independent dimensions:

\begin{itemize}
    \item \textbf{Bandwidth / RTT}: 18 paired combinations covering realistic
          link types from 50\,Mbit/s mobile (30--80\,ms RTT) through 5\,Gbit/s
          datacenter (10--150\,ms RTT), including high-BDP scenarios where the
          bandwidth-delay product exceeds the 32\,MB per-stream receive window
          configured for msquic.
    \item \textbf{Packet loss}: 0\%, 0.1\%, 0.5\%, 1\%, 2\%.
    \item \textbf{Jitter}: 0, 2, 5, 10\,ms (uniform random, applied per packet).
    \item \textbf{Corruption}: 0\%, 0.05\%, 0.1\%, 0.5\%.
    \item \textbf{Competing traffic}: none, TCP at 40\% of link bandwidth,
          UDP at 40\% of link bandwidth.
\end{itemize}

The Cartesian product of these parameter sets yields 4,320 unique network condition combinations.
Each combination is repeated 10 times per implementation to average out
congestion-control variability, resulting in 43,200 runs per implementation
and 216,000 total runs across the five implementations. Each run transfers data
continuously for 30 seconds.

\subsection{Execution environment and scale}
\label{subsec:execution}

Initial development and validation were performed on WSL2. In practice,
WSL2's partial kernel support caused recurring failures: certain kernel
modules required by the implementations under test~\cite{cloudflare-quiche}
could not be loaded, and network namespace operations became unreliable at
high bandwidths. Therefore, testing was migrated to a native Linux environment,
where all kernel features were available.

The full dataset was collected on the faculty computing cluster over
approximately 12 days. Sixteen virtual machines were provisioned on an
OpenStack deployment, each with the \texttt{c.medium} flavour: 4 vCPUs,
4\,GB of RAM, and a 25\,GB root volume. An additional 25\,GB block volume
was attached to each instance to avoid running out of disk space due to packet
captures. During development, unfiltered pcap files occasionally grew
very large, so the extra storage was added as a precaution. After
tightening the capture filters and scoping the scenarios, disk usage
remained modest and the additional volumes were ultimately unnecessary,
though they were kept in place.

The 43,200 runs per implementation were distributed across the 16 machines
using a deterministic sharding scheme (\texttt{--shard K --of 16}).
Each machine processed an independent, non-overlapping subset of the
configuration matrix, and results were consolidated after the runs completed.

\subsection{Implementation configuration and parity}
\label{subsec:config-parity}

The goal is a fair comparison of each implementation as its own documentation
presents it for a bulk-throughput workload, not an artificial equalisation. In
practice, three things are applied uniformly across all five stacks. First,
each is given a flow-control (or receive-buffer) window scaled to the
bandwidth-delay product, the QUIC analogue of sizing the TCP receive buffer,
since the stacks' out-of-the-box windows are too small to fill a high-BDP pipe.
Second, every container applies the same OS-level socket-buffer tuning
(\texttt{net.core} \texttt{rmem}/\texttt{wmem} raised to 64\,MB) and uses CUBIC
congestion control. Third, each stack uses the high-throughput send path its
own documentation recommends, namely Generic Segmentation Offload and kernel
pacing. For quiche this is written in application code following Cloudflare's
reference apps~\cite{cloudflare-quiche-gso}, because quiche is a sans-I/O
library that leaves the socket to the caller. In contrast, msquic, mvfst, and
quic-go provide the same mechanisms internally and only need to be enabled.

Two configuration details are worth recording. quic-go v0.47 aborts connections
above roughly 1\,Gbit/s with a "too many gaps" error, owing to a hard-coded
limit of 1000 stream-frame gaps. Due to this reason, the limit is raised to 100000
by a one-line source patch so the connection survives at the bandwidths tested.
And while msquic receives the same BDP-scaled windows as the others, it still
degrades at high bandwidth-delay products (Section~\ref{subsec:fc-limited}).
That behaviour is reported as measured rather than tuned around.

% =============================================================================
\section{Dataset Construction}
\label{sec:dataset}

Each completed run produces a result record containing the network condition
parameters that were configured, the implementation that was tested, and a set
of measured performance metrics. These raw records are aggregated and labelled
by \texttt{train.py} before being used for model training.

\subsection{Measured metrics}
\label{subsec:metrics}

The following performance metrics are recorded per run:

\begin{itemize}
    \item \textbf{goodput\_mbps}: application-layer throughput, excluding
          QUIC and transport headers.
    \item \textbf{overhead\_pct}: ratio of protocol overhead to total bytes
          on the wire, expressed as a percentage.
    \item \textbf{hs\_ms}: handshake duration in milliseconds, measured
          from the first Initial packet to the completion of the handshake.
    \item \textbf{retrans\_pct}: percentage of packets retransmitted.
    \item \textbf{recovery\_s}: time in seconds until throughput recovers
          after a loss event.
    \item \textbf{tail\_rtt\_ms}: 99th-percentile round-trip time observed
          during the transfer.
    \item \textbf{srv\_cpu\_mean\_pct}, \textbf{cli\_cpu\_mean\_pct}: mean
          CPU usage on the server and client respectively.
    \item \textbf{srv\_mem\_mean\_mib}, \textbf{cli\_mem\_mean\_mib}: mean
          memory usage on the server and client respectively.
\end{itemize}

Runs where the transfer does not complete within the 30-second measurement window
(due to a failed handshake, mid-transfer stall or timeout) are
flagged as \texttt{timed\_out}. These rows are retained in the dataset for
Stage~1 (timeout classifier training) but excluded from Stage~2 label
assignment, since a timed-out run produces no meaningful performance metrics.

A data quality filter removes mvfst rows where the measured handshake time
exceeds 10,000\,ms. These represent occasional fake outliers in which
mvfst entered an unusually long retry cycle, producing a handshake time
several orders of magnitude above its normal range. Replacing them with NaN
would propagate into the per-condition averages and distort the scoring.
Dropping them entirely is the safer choice.

\subsection{Deployment profiles and composite scoring}
\label{subsec:profiles}

Rather than training a single model that predicts a globally best
implementation, the recommender is built around a generic scoring mechanism:
any deployment scenario can be expressed as a named set of (metric, direction,
weight) triples and a separate Stage~2 classifier is trained for each set
of this kind. For each profile, a scalar composite score is computed per
(condition, implementation) pair as a weighted sum of normalised metrics and
the implementation with the highest score within a condition becomes the
training label for that profile.

This design means that adding a new profile requires only defining its weight
vector in \texttt{train.py} and rerunning Stage~2 training. The rest of the
pipeline (normalisation, label assignment, cross-validation, model serialisation)
is driven by the same loop and requires no further changes.

Five profiles were defined as examples of realistic deployment scenarios:

\begin{itemize}
    \item \textbf{Datacenter}: maximise throughput above all else.
          Goodput carries a weight of 0.80, thus penalties are small for
          overhead, retransmissions, tail latency, and server CPU.
    \item \textbf{Lightweight}: optimise for resource efficiency on
          constrained hardware. CPU and memory weights together account for
          0.55 of the score, while goodput is weighted at only 0.25.
    \item \textbf{Mobile lossy}: prioritise resilience to packet loss.
          Retransmission rate (0.25) and loss recovery time (0.15) dominate
          alongside goodput (0.40).
    \item \textbf{Latency sensitive}: minimise perceived delay. Tail RTT
          (0.30) and handshake time (0.20) are weighted heavily alongside
          goodput (0.30).
    \item \textbf{Deep space}: optimise for extreme-RTT, high-loss links.
          Retransmission rate (0.25), recovery time (0.15), and tail RTT (0.15)
          together outweigh goodput (0.30), with additional penalties for
          overhead and handshake time.
\end{itemize}

An operator with different priorities, a video streaming service that
cares primarily about tail latency at moderate loss rates for example,
could define a custom profile with appropriate weights and obtain a tailored
classifier from the same dataset without re-running any experiments.

\subsection{Per-condition normalisation}
\label{subsec:normalisation}

A critical design decision is that metric normalisation is performed
\textit{within each condition} rather than globally. For each network
condition, each metric is scaled to $[0, 1]$ relative to the minimum and
maximum values observed across the five implementations under that exact
condition. This means the composite score reflects how much better an
implementation performs relative to the best and worst alternatives available
in the same environment.

Global normalisation would be incorrect here: TCP achieves approximately
4,000\,Mbps goodput in the highest-bandwidth datacenter conditions, which would
compress all other conditions into a narrow band near zero after global scaling.
Under global normalisation, the 0.05-weight overhead penalty would then
numerically dominate the 0.80-weight goodput term for most conditions, producing
winners driven by overhead rather than throughput.

\subsection{Derived feature: fc\_limited}
\label{subsec:fc-limited}

A binary feature \texttt{fc\_limited} is derived from the bandwidth and RTT
values of each condition. For msquic (v2.4.4) we configure the initial
per-stream receive window at 32\,MB and the connection-level flow-control
window at 128\,MB, overriding msquic's built-in defaults~\cite{microsoft-msquic}
of 64\,KB (\texttt{QUIC\_DEFAULT\_STREAM\_FC\_WINDOW\_SIZE}) and 16\,MB
(\texttt{QUIC\_DEFAULT\_CONN\_FLOW\_CONTROL\_WINDOW}). From this initial value
msquic auto-tunes the per-stream window upward, doubling it on an RTT-gated
schedule and capping it at the connection-level window.

Empirically, msquic's reliability drops sharply once the bandwidth-delay
product exceeds the 32\,MB initial window: it times out in 32.3\% of conditions
where \texttt{fc\_limited=1}, compared to 6.7\% elsewhere, rising to 44.4\% at
RTT\,=\,300\,ms. The \texttt{fc\_limited} feature marks this regime:

\[
    \texttt{fc\_limited} = \left[ \frac{\texttt{bandwidth\_mbps} \times
    \texttt{rtt\_ms}}{8000} > 32 \right]
\]

We treat the 32\,MB boundary as an empirically determined threshold rather than
a hard protocol limit. Including \texttt{fc\_limited} explicitly gives the
Random Forest a concrete tell for this regime rather than requiring it to
discover the threshold implicitly from the continuous bandwidth and RTT
features alone.

% =============================================================================
\section{Training Pipeline}
\label{sec:training}

Model training is implemented in \texttt{train.py} as a two-stage pipeline.
Stage~1 identifies implementations that are likely to time out under the given
conditions and eliminates them from consideration. Stage~2 ranks the surviving
implementations according to the deployment profile. Training both stages from
the same dataset means the two classifiers share a consistent view of what
"good" and "bad" conditions look like.

\subsection{Cross-validation strategy}
\label{subsec:cv}

All models are evaluated with 5-fold GroupKFold cross-validation, where the
grouping key is the condition identifier, a unique integer assigned to each
distinct (bandwidth, RTT, loss, jitter, corruption, competing) tuple. This
ensures that all 10 repetitions of a given condition are placed in the same
fold and never split across training and test sets.

Without this constraint, repetitions of the same condition would appear on
both sides of a fold boundary. Because repetitions differ only in random
CUBIC cycle variation rather than in any predictable pattern, a model evaluated
this way would appear more accurate than it truly is: it would effectively
be tested on conditions it has already seen in training.

\subsection{Stage 1: timeout classifiers}
\label{subsec:stage1}

Five binary classifiers are trained, one per implementation. Each classifier
receives the eight network features and predicts whether a run under those
conditions will time out. The Random Forest is configured as follows:

\begin{verbatim}
RandomForestClassifier(
    n_estimators=300,
    max_features='sqrt',
    min_samples_leaf=5,
    class_weight='balanced',
    oob_score=True,
)
\end{verbatim}

The \texttt{class\_weight='balanced'} setting is necessary because timeouts
are rare for most implementations: only msquic produces a meaningful number
of timed-out rows (5,363 out of 43,200), while quic-go has zero. Without
balancing, a classifier that always predicts "no timeout" would achieve
over 99\% accuracy while being useless for the task. Balanced weighting
up-samples the minority class. This prevents timeouts from being ignored
during training where they would otherwise be overwhelmed by non-timeout
samples.

At inference time, Stage~1 computes a timeout probability for each
implementation. Any implementation whose probability exceeds a configurable
threshold (default 10\%) is eliminated before Stage~2 runs. If all
implementations are eliminated, which can occur at extreme conditions,
the one with the lowest timeout probability is kept as a fallback.

\subsection{Stage 2: profile classifiers}
\label{subsec:stage2}

Five multiclass classifiers are trained, one per deployment profile. Each
classifier is trained on the averaged, per-condition normalised dataset
described in Section~\ref{sec:dataset}: one row per (condition, implementation)
pair, labelled with the winner under that profile's composite score. The
configuration is:

\begin{verbatim}
RandomForestClassifier(
    n_estimators=300,
    max_features='sqrt',
    min_samples_leaf=3,
    class_weight='balanced',
    oob_score=True,
)
\end{verbatim}

The lower \texttt{min\_samples\_leaf} relative to Stage~1 (3 vs.\ 5) allows
the Stage~2 trees to capture finer-grained distinctions between closely ranked
implementations, at the cost of slightly higher variance. This trade-off is
appropriate here because Stage~2 labels are already averaged over repetitions
and are therefore less noisy than the raw per-run Stage~1 labels.

\subsection{Feature importance}
\label{subsec:feature-importance}

Two complementary importance measures are computed for each Stage~2 model.
Mean Decrease in Impurity (MDI) is read directly from the final fitted model
and reflects how much each feature reduces Gini impurity across all trees.
Permutation importance is computed on the last cross-validation fold's held-out
set: features are shuffled one at a time and the resulting accuracy drop is
recorded. Because permutation importance is measured on data the model has
never seen during training, it gives a less optimistic estimate of each
feature's true contribution than MDI alone.

Both measures are saved to \texttt{results/feature\_importance.csv} alongside
the cross-validation accuracy and out-of-bag scores for each profile.

% =============================================================================
\section{Inference Module}
\label{sec:inference}

The trained models are exposed through \texttt{infer.py}, which can be used
both as a Python library and as a standalone command-line tool. The public
entry point is a single \texttt{recommend()} function that accepts a
dictionary of network condition values and a profile name, runs both stages
of the pipeline, and returns a structured result.

\subsection{Inference flow}
\label{subsec:inference-flow}

Given a set of network conditions, inference proceeds as follows:

\textbf{Pre-processing.} If \texttt{fc\_limited} is absent from the input,
it is computed automatically from bandwidth and RTT using the same formula
used during training (Section~\ref{subsec:fc-limited}). This means a caller
only needs to supply the seven directly observable features, while the derived
feature is handled transparently.

\textbf{Stage 1.} Each of the five per-implementation timeout classifiers
is queried. For each, \texttt{predict\_proba} returns the probability of
a timeout. Implementations whose probability exceeds the elimination
threshold (default 10\%) are marked as eliminated. If all implementations
are eliminated, the one with the lowest timeout probability is kept as a
fallback to ensure a recommendation is always produced.

\textbf{Stage 2.} The profile classifier for the requested deployment profile
is queried with the same feature vector. It returns a probability for each
implementation. Probabilities belonging to eliminated implementations are
discarded, and the remaining values are renormalised to sum to 1. The
implementation with the highest renormalised probability is returned as
the top recommendation.

The renormalisation step is required because otherwise the confidence reported
to the user would be lessened by the probability mass of removed implementations,
making strong recommendations appear weaker than they truly are.

\subsection{Model caching and library interface}
\label{subsec:model-caching}

Each \texttt{.pkl} model file is loaded with \texttt{joblib.load} on first
use and stored in a module-level dictionary. Subsequent calls to
\texttt{recommend()} with any profile or implementation reuse the cached
objects, so the I/O cost of deserialising ten model files is paid only once
per process lifetime. This matters in the network prober, which calls
\texttt{recommend()} multiple times per round across multiple rounds.

The function returns a dictionary containing the full ranked list with
renormalised probabilities, the top implementation, the confidence value,
the list of eliminated implementations, the raw per-implementation timeout
probabilities, and a flag warning when \texttt{fc\_limited=1} and msquic
has not been eliminated (a condition where msquic is at elevated timeout
risk even if its Stage~1 probability falls just below the threshold).

% =============================================================================
\section{Network Prober}
\label{sec:prober}

The network prober closes the loop between measurement and recommendation.
It consists of two components: \texttt{probe\_server.py}, a lightweight
daemon deployed on the remote host, and \texttt{prober.py}, the client-side
tool that measures network conditions and calls \texttt{recommend()}.

\subsection{Cooperative and passive modes}
\label{subsec:probe-modes}

The prober supports two operating modes depending on whether
\texttt{probe\_server.py} is running on the target host.

\textbf{Cooperative mode} is enabled by passing \texttt{--server-port}.
The prober runs three measurements in sequence: ping for RTT, loss, and
jitter respectively, a UDP iperf3 test for bandwidth and competing traffic
and a UDP CRC32 burst for corruption. The sequencing is deliberate: ping
runs first because it has no side effects, iperf3 runs second because it
generates significant traffic that would inflate the competing-traffic
estimate if measured concurrently, and the corruption burst runs last
because it requires the probe server.

\textbf{Passive mode} is used when no probe server is available. Ping is
still run for RTT, loss, and jitter, and competing traffic is estimated
from a one-second NIC counter window via \texttt{psutil}. Bandwidth cannot
be measured passively and must be supplied by the operator via
\texttt{--bw}. Corruption defaults to zero. This mode is appropriate when
the target host is not under the operator's control.

\subsection{Bandwidth measurement}
\label{subsec:bw-measurement}

Bandwidth is measured using iperf3 in UDP mode (\texttt{-u -b 10G}).
The sender pushes at 10\,Gbit/s, the link rate is the bottleneck. The
received bandwidth is derived from the sender rate and the loss fraction
reported by the iperf3 receiver:

\[
    \texttt{bandwidth\_mbps} = \frac{\texttt{bits\_per\_second} \times
    (1 - \texttt{lost\_percent} / 100)}{10^6}
\]

TCP iperf3 was considered but rejected: TCP goodput collapses under loss
because congestion control interprets loss as a congestion signal and
reduces the sending rate. This would cause the prober to report the
achieved throughput rather than the link capacity, which is the feature
the model was trained on.

Competing traffic is derived from the same iperf3 window: the NIC byte
counters are sampled before and after the iperf3 test, and any traffic
beyond what iperf3 itself sent is attributed to background load.

\subsection{Corruption measurement}
\label{subsec:corruption-measurement}

\texttt{probe\_server.py} runs a UDP echo server on a configurable port.
Each probe packet carries a 4-byte sequence number, a 1000-byte payload
of known content, and a 4-byte CRC32 checksum of the payload. The server
verifies the checksum and replies with the sequence number and a boolean
indicating whether corruption was detected.

The prober sends all 1,000 probe packets in a burst first, then collects
replies until a deadline. Sending and receiving are decoupled so that the
total probe time is bounded by one round-trip worth of latency rather than
scaling with the number of packets. Early exit is applied when all replies
have been received, otherwise collection continues until the deadline,
since gaps between replies are normal on congested links.

At 1,000 packets, the probe reliably detects corruption rates down to
approximately 0.1\% (1 expected corrupted packet). Lower rates require
more packets to detect with statistical confidence.

\subsection{Stability averaging}
\label{subsec:stability}

A single probe round can be misleading on a variable link. The prober
runs $N$ rounds (default 3) and averages the \texttt{recommend()} output
across rounds. Probabilities from each round's ranking are averaged and
renormalised, timeout probabilities are averaged independently and
eliminated implementations are re-derived from the averaged timeout
probabilities. This ensures that an implementation eliminated in most
rounds but not the last still appears eliminated in the final output.

The confidence of the averaged recommendation is compared against a
stability threshold (default 60\%). If confidence falls below this value,
the output notes that the top two implementations are sufficiently close
that either would be a reasonable choice, and suggests running more rounds.

\subsection{StableRecommender: hysteresis for autonomous deployment}
\label{subsec:stable-recommender}

For scenarios where the prober runs continuously and drives automatic
re-deployment, a \texttt{StableRecommender} class provides a hysteresis
gate. It tracks a candidate challenger to the currently deployed
implementation and only triggers a switch when the challenger has been
top-ranked for a configurable number of consecutive rounds (default 3)
\textit{and} its probability gap over the current implementation exceeds
a minimum margin (default 0.15). This prevents thrashing between
near-equivalent implementations on noisy links, where small fluctuations
in measured conditions would otherwise cause unnecessary redeployments.
The class is not wired to the CLI and is not exercised in the evaluation,
yet it is included as a proof-of-concept interface for future autonomous
deployment scenarios.
