\chapter{Introduction}
\label{chapter:introduction}

QUIC is a UDP-based transport protocol that now carries a large share of web
traffic, and unlike TCP it is implemented entirely in user space. That has led
to many independent implementations, written by different organisations in
different languages, all following the same specification but not behaving the
same way. The differences are not small. Jaeger et al.~\cite{jaeger2023quic}
benchmarked several QUIC stacks on high-rate links and found that throughput
under the same network conditions ranged from under 100\,Mbit/s to nearly
5\,Gbit/s depending only on which client and server implementation was used.
QUIC is also not automatically faster than TCP: Zhang et
al.~\cite{zhang2024quic} measured the QUIC stack losing up to 45\% of its
throughput against TCP on fast connections, because of the extra per-packet CPU
cost of running in user space.

So the question "which transport should I deploy?" does not have a fixed
answer. The best choice depends on the network the connection will run over and
on what the deployment actually cares about, whether that is raw throughput, low
latency, resilience to packet loss, or CPU efficiency on constrained hardware.
Today that decision is made by hand, if it is made deliberately at all: an
engineer picks an implementation based on familiarity or reputation and rarely
revisits it. There is no tool that takes measured network conditions and turns
them into a recommendation. Closing that gap is the goal of this thesis.

% =============================================================================
\section{Research Questions}
\label{sec:intro-rqs}

This work is organised around three research questions.

\textbf{RQ1: Does the best transport implementation actually depend on network
conditions and deployment priorities, or is there a single overall winner?}
Before building anything that recommends an implementation, it is worth
establishing that the choice is real. This question is answered empirically, by
measuring all five implementations across a wide grid of network conditions and
checking whether the winner shifts as bandwidth, RTT, loss, and the optimisation
target change.

\textbf{RQ2: Can a model trained offline on emulated measurements predict the
right implementation from observable network features, and how accurately?}
If the choice depends on conditions, the next question is whether it can be
learned. This means training a model on the collected dataset, measuring how
well it generalises to conditions it has not seen, and being honest about where
and why it fails.

\textbf{RQ3: Can such a recommender run against a live connection?} A model that
only works on a pre-recorded dataset is of limited use. The final question is
whether the same model can be driven by a tool that measures a real connection's
conditions on the spot and produces a usable recommendation, together with a
signal of how confident it is.

% =============================================================================
\section{Contributions}
\label{sec:intro-contributions}

This thesis makes the following contributions.

\textbf{A custom benchmarking framework and dataset.} A three-container Docker
setup applies reproducible network conditions with \texttt{tc/netem} and
measures five transport implementations (quiche, msquic, mvfst, quic-go, and a
TCP\,+\,TLS\,1.3 baseline). Running it across a systematic grid of conditions
produced a dataset of 216{,}000 labelled runs.

\textbf{A two-stage recommender.} The model first eliminates implementations
likely to time out under the given conditions, then ranks the survivors with a
profile-specific classifier built on random forests. The profile scoring is
generic: a new deployment priority can be added by defining a weight vector and
retraining only Stage~2 on the existing dataset, without collecting any new
measurements. Extending the dataset itself, with either a new implementation or with
new conditions, is a separate offline step (see Section~\ref{subsec:normalisation}).

\textbf{A live network prober.} A deployment tool measures a real connection's
conditions and feeds them to the recommender, returning a ranked recommendation
with a confidence value and a stability flag, in both a cooperative and a
passive measurement mode.

\textbf{An empirical study of the result.} The evaluation shows that the best
implementation genuinely shifts with conditions and deployment profile, reports
the recommender's accuracy per stage and per profile, and accounts honestly for
its limits, in particular the cases where two implementations are too close to
separate reliably.

% =============================================================================
\section{Scope of the Benchmark}
\label{sec:intro-scope}

One point needs stating up front, because it shapes how every later result
should be read. This is not an apples-to-apples comparison of identically
configured stacks. Each implementation is configured the way its own
documentation presents it for a bulk-throughput workload, and where the
documentation says nothing, it is left as it ships. Concretely, every stack is
given a flow-control window scaled to the bandwidth-delay product (the QUIC
analogue of sizing the TCP receive buffer), the same operating-system socket
buffer tuning, CUBIC congestion control, and whatever high-throughput send path
the project documents, such as Generic Segmentation Offload and kernel pacing.
Where a stack requires the application to wire this up by hand, as quiche does
because it leaves the socket to the caller, that work is done following the
project's own reference code. Consequently, where a stack provides it internally, it is simply
enabled. Section~\ref{subsec:config-parity} gives the details.

The consequence is that the numbers in this thesis describe each implementation
as a developer following its documentation would experience it, not some
normalised lower bound, and not a tuning contest either. Differences that come
from one project shipping an optimisation another lacks are treated as real,
because a developer choosing between them would feel exactly those differences.
The measurements also come from controlled emulation on a single Linux host
rather than the open internet, and focus on single-stream bulk transfer rather
than web page loads, so the goal throughout is repeatable relative comparison,
not absolute production figures.

% =============================================================================
\section{Thesis Organisation}
\label{sec:intro-organisation}

Chapter~\ref{chapter:background} introduces the concepts the rest of the thesis
relies on: QUIC and why it exists, the implementations under test, the network
parameters, the emulation tooling, and random forests.
Chapter~\ref{sec:architecture} describes how the data-collection environment,
the training pipeline, and the live prober fit together.
Chapter~\ref{chapter:implementation} covers the implementation in detail, from
the container topology and condition injection through to the training pipeline
and the prober. Chapter~\ref{chapter:evaluation} presents the results: raw
performance, the profile label distributions, model accuracy, feature
importance, and the live prober validation. Chapter~\ref{chapter:rel-work}
positions the work against prior QUIC studies, and
Chapter~\ref{chapter:conclusion} revisits the research questions and concludes.
