\chapter{Related Work}
\label{chapter:rel-work}

QUIC has been studied from several angles: measuring how it performs, changing
it to perform better, documenting how much its implementations differ, and
applying machine learning to it. This thesis sits next to all four but does
something none of them do, which is to recommend which existing implementation
to deploy for a given set of network conditions. This chapter goes through the
closest work in each group and says where this thesis differs.

% =============================================================================
\section{QUIC: Existing Studies}
\label{sec:rw-studies}

% -----------------------------------------------------------------------------
\subsection{Performance Evaluation and Benchmarking}
\label{subsec:rw-benchmarking}

The most directly related line of work measures QUIC throughput and latency
under controlled conditions. Jaeger et al.~\cite{jaeger2023quic} extended the
QUIC Interop Runner to benchmark several implementations on 10\,Gbit/s links
and found that achieved throughput ranged from under 100\,Mbit/s to nearly
5\,Gbit/s depending only on which client and server implementation was used,
and that default OS buffer sizes were a common bottleneck. The journal version
of that work~\cite{kempf2024quic} extends the same setup with more
implementations and conditions. This is the closest study in spirit to the
data collection here: it shows clearly that the choice of implementation
dominates performance on fast links. The difference is that it stops at
reporting the numbers, whereas this thesis \emph{turns the same kind of
measurements into an automatic recommendation}.

Zhang et al.~\cite{zhang2024quic} looked at the opposite end and showed that
the QUIC stack can lose up to 45\% of its throughput against TCP on very fast
connections, mostly because of the extra per-packet CPU cost of running in
user space. Yu and Benson~\cite{yu2021dissecting} measured QUIC against
production endpoints from Google, Facebook, and Cloudflare across a range of
network conditions, workloads, and client implementations, and also found wide
variation between providers. Shreedhar et al.~\cite{shreedhar2022evaluating}
compared QUIC to TLS/TCP over web, cloud storage, and video workloads from
several vantage points, including a low-bandwidth high-RTT residential link.
All three confirm that QUIC sometimes wins and sometimes loses, and that the
result depends heavily on the conditions, which is exactly the premise this
thesis builds on. However, none of them frame the result as a decision to be
made automatically. Instead, they \emph{describe the landscape rather than
navigate it}.

A more specialised case is Dizdarevic and Jukan~\cite{dizdarevic2021benchmarking},
who benchmarked HTTP/QUIC in an IoT cloud/edge setting. It is mentioned here
only briefly, as it fixes a single deployment scenario rather than sweeping
conditions, but it is another example of the same observation that QUIC's
benefit is conditional on the environment.

% -----------------------------------------------------------------------------
\subsection{Improving QUIC Internals}
\label{subsec:rw-improving}

A separate group of papers tries to make QUIC itself faster rather than
compare existing versions. Kharat and Kulkarni proposed a modified QUIC that
changes the handshake and the way the congestion window is updated to cut
control overhead, first without~\cite{kharat2019modquic} and then
with~\cite{kharat2021modquic} a reworked congestion control, reporting
throughput gains over both stock QUIC and TCP. Engelbart and
Ott~\cite{engelbart2021rtp} went after a specific use case, congestion control
for real-time media carried over QUIC. These efforts are orthogonal to this
thesis: they modify one implementation to raise its ceiling, while this work
\emph{takes the implementations as they ship and picks the best one for the
conditions}. A better congestion controller would change which implementation
wins, but it would not remove the need to choose.

% =============================================================================
\section{Implementation Diversity}
\label{sec:rw-diversity}

The assumption underneath this entire thesis, that different QUIC
implementations behave differently even though they follow the same
specification, was studied directly by Marx et al.~\cite{marx2020same}. They
compared several stacks and found large differences in flow control,
congestion control, and stream multiplexing, along with outright bugs and
missing optimisations. Their conclusion is that "which QUIC" is a real
question with real performance consequences. This thesis takes that conclusion
as its starting point. Where Marx et al. document that the choice matters and
leave it to the reader to act on it, the recommender here \emph{turns that
choice into something a system can make on its own from measured conditions}.

% =============================================================================
\section{Machine Learning for QUIC and Network-Aware Selection}
\label{sec:rw-ml}

Machine learning has been applied to QUIC, but for different tasks.
Almuhammadi et al.~\cite{app2023quicensemble} used ensemble methods, including
random forests, to classify QUIC network traffic, which is a useful precedent
for random forests working well on QUIC-related tabular data, but the goal is
identifying traffic rather than choosing an implementation. Zhang et
al.~\cite{zhang2023pbq} went further inside the protocol with PBQ, a deep
reinforcement learning congestion controller for QUIC that outperformed Cubic
and BBR on throughput and RTT. That is machine learning tuning one
implementation online, which is a different problem from selecting among
several offline, and it is the kind of explore-and-exploit approach discussed
and set aside in Section~\ref{sec:bg-random-forest}.

This is the gap the thesis fills. Prior work measures QUIC, improves a single
implementation, documents that implementations differ, or applies machine
learning to QUIC for classification and congestion control. \textbf{None of it
builds a network-aware model that recommends which existing implementation to
deploy.} That recommendation, learned offline from a systematic sweep of
conditions across five implementations, is the contribution of this work.
